### 2016-08-29 notions.tex
privacy definitions

\section{Security Notions}
\label{sec:security-notions}
Following NY, we are interested in defining an adaptive (i.e.,
adversarial) notion of correctness.
In addition to correctness, we establish three notions of privacy for
set-representations. The first two have the flavor of a one-wayness
game, in which the adversary seeks to return an element of the
represented set~$S$.  The final notion is reminiscent of
semantic-security for encryption schemes, and attempts to capture the
intuition that the representation~$M$ and the public
information~$\pubaux$ should not leak any efficiently computable
information about~$S$.

@@

\subsection{Notions of Privacy}
One can also consider the orthogonal notion of privacy for a
data structure. \jnote{I am still hoping that we can treat data structures
at an even more general level than set-representation data structures.}
We formalize two types of privacy notions. The first captures
a ``one-wayness'' property, which (informally) requires that an attacker
cannot learn any value in a set $S$ given its
representation $M$ along with the public information~$\pub$.  The second definition
captures a flavor of semantic security, essentially requiring that $M$ and $\pub$ do
not leak any (non-trivial) information about the set~$S$ they represent.
%\jnote{Actually, I am now unsure what the right way to define it is. It seems natural
%to give the attacker $\aux$. But if we give them $\aux$ and $M$ then they have everything.
%If $S$ has high min-entropy (element-wise), then it might still be
%possible to have security here.}\tsnote{I think we will always bump up
%against the (element-wise) min-entropy of the distribution over~$S$,
%at least in the ``one-wayness'' style definition that I had been imagining. }

\jnote{OK, from here on I specialize to set-representation data structures\ldots}
The one-wayness definition (see
Figure~\ref{fig:privacy-ow}) is parameterized by a distribution $\distr{}{}$ over
sets. \jnote{I got rid of the weaker definition, since it seems unnatural to me.}
A set $S$ is chosen from this distribution, and a representation
$(M,\pubaux,\privaux)$ of that set is computed.
In one experiment ($\ExpPriv{\Pi,\distr{}{}}$), the attacker is given $M, \pubaux$;
in a second experiment ($\ExpPrivNull{\Pi,\distr{}{}}{}$), it is not.
In both cases, the attacker is given
access to an oracle via which it can
issue queries and obtain the corresponding responses.
For a given $\distr{}{}$ and $A$, we define the advantage measure
$\AdvPriv{\Pi,\distr{}{}}{A} \bydef
\left| \Prob{\ExpPriv{\Pi,\distr{}{}}{A}=1} - \Prob{\ExpPrivNull{\Pi,\distr{}{}}{A}=1} \right|$.
We track the
time-complexity~$t$ and number of oracle queries~$q$ made by the adversary.

\begin{figure}[hbtp]
\centering
\fpage{.6}{
\hpagess{.5}{.45}
{
\experimentv{$\ExpPriv{\setprim,\distr{}{}}{A}$}\\
$S \getsr \distr{}{}$\\
$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
$z \getsr A^{\TestOracle}(M,\pubaux)$\\
if $z \in S$ then return 1\\
return 0\\

\medskip
\experimentv{$\ExpPrivNull{\Pi, \distr{}{}}{A}$} \\
$S \getsr \distr{}{}$\\
$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
$z \getsr A^{\TestOracle}(\emptystring,\emptystring)$\\
if $z \in S$ then return 1\\
return 0\\
%\medskip
%\experimentv{$\ExpWPriv{\setprim,\distr{}{}}{A}$}\\
%$S \getsr \distr{}{}$\\
%$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
%$z \getsr A^{\TestOracle}(\pubaux)$\\
%if $z \in S$ then Return 1\\
%Return 0\\
}
%
{
\oracle{$\TestOracle(q)$}\\
$a \gets \Qry(M,\pubaux,\privaux,q)$\\
return~$a$
}
}
\caption{Experiments for a ``one-wayness'' style privacy
  definition.}
\label{fig:privacy-ow}
\end{figure}

\jnote{I'm not sure whether we want to limit consideration to ``high-entropy''
distributions
and require the probability of finding an element to be low,
or whether we want to allow arbitrary distributions but compare the attacker's
success probability when given $(M, \pub)$ to its success probability when given nothing.
The latter seems cleaner, so I took that approach here.}

\jnote{Actually, I am unsure whether a one-wayness definition is independently
interesting if we have (and can achieve) a semantic-security definition.}

Our semantic-secure definition is simulation-based (see
Figure~\ref{fig:privacy-ss}).
In this definition we compare the probabilities with which an attacker $A$ outputs~1
in two experiments: one ``real'' and one ``simulated.''
In the real experiment (ss-rep-1), $A$ outputs a set $S$ and is given the
$M,\pubaux$ that result from running $\Rep(S)$.
In the simulated experiment (ss-rep-0), $A$ outputs $S$ and is given $M, \pubaux$ output
by a simulator $\calS$ that is only given ${\sf leak}(S)$. \jnote{Note
there is no need for a query oracle here, since the attacker knows the set. Interestingly,
if we provide a query oracle in the real world and an oracle that returns the correct
answer in the simulated world, then this definition seems to imply
correctness also.}
\tsnote{Need to pass state between stages if you want~$A$ to know the
  set it previously sampled, right?}
For a given function ${\sf leak}$, adversary $A$, and simulator~$\calS$,
we define the advantage measure
$\AdvPrivSS{\Pi,{\sf leak}}{A,\calS} \bydef
\left| \Prob{\ExpPrivSSreal{\Pi}{A}=1} -
\Prob{\ExpPrivSSsim{\Pi,{\sf leak}}{A, \calS}=1} \right|$.
Informally, a scheme $\Pi$ is secure with respect to some leakage ${\sf leak}$
if there is a simulator $\calS$ such that $\AdvPrivSS{\Pi,{\sf leak}}{A,\calS}$ is small
for all~$A$. Intuitively, this means that the public portion
$M, \pubaux$ of the representation of a set~$S$ leaks only ${\sf leak}(S)$.

\begin{figure}[hbtp]
\centering
\fpage{.6}{
\hpagess{.5}{.45}
{
\experimentv{$\ExpPrivSSreal{\Pi}{A}$}\\
$S \getsr A()$ \\
$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
return $A(M,\pubaux)$\\
}
%
{
\experimentv{$\ExpPrivSSsim{\Pi, {\sf leak}}{A, \calS}$}\\
$S \getsr A()$\\
$(M,\pubaux) \getsr \calS({\sf leak}(S))$ \\
return $A(M,\pubaux)$\\

\medskip
\experimentv{$\ExpPrivSSsim{\Pi}{A, \calS}$}\\
$S \gets A()$\\
$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
$(M',\pubaux) \gets \calS(\pubaux)$ \\
return $A(M',\pubaux)$\\
}
}
\caption{Experiments for a ``semantic-security'' style privacy
  definition.}
\label{fig:privacy-ss}
\end{figure}

>>>>

\section{Security Notions}
\label{sec:security-notions}
Following NY, we are interested in defining an adaptive (i.e.,
adversarial) notion of correctness.
In addition to correctness, we establish notions of privacy for
set-representations.

@@

\subsection{Notions of Privacy}
One can also consider the orthogonal notion of privacy for a
data structure. \jnote{I am still hoping that we can treat data structures
at an even more general level than set-representation data structures.}
We formalize two types of privacy notions. The first captures
a ``one-wayness'' property that (informally) requires that
the representation of some set~$S$ does not make it any easier for an attacker
to learn an element of~$S$.
The second definition
captures a flavor of semantic security, essentially requiring that $M$ and $\pub$ do
not leak any information about the set~$S$ they represent beyond
what is captured in some explicit leakage function.

\jnote{OK, from here on I specialize to set-representation data structures\ldots}

\jnote{I don't know if I like the one-wayness definition any more; it seems a bit
ad hoc. And why do we care about one-wayness, anyway? Instead, I give
two versions of semantic security.}

%The one-wayness definition (see
%Figure~\ref{fig:privacy-ow}) is parameterized by a distribution $\distr{}{}$ over
%sets. A set $S$ is chosen from this distribution, and a representation
%$(\pubaux,\privaux)$ of that set is computed.
%In one experiment ($\ExpPriv{\Pi,\distr{}{}}$), an attacker~$A$ is given $\pubaux$;
%in a second experiment ($\ExpPrivNull{\Pi,\distr{}{}}{}$), it is not.
%For a given $\distr{}{}$ and $A$, we define the advantage measure
%$\AdvPriv{\Pi,\distr{}{}}{A} \bydef
%\left| \Prob{\ExpPriv{\Pi,\distr{}{}}{A}=1} - \Prob{\ExpPrivNull{\Pi,\distr{}{}}{P}=1} \right|$.
%We track the
%time-complexity~$t$ and number of oracle queries~$q$ made by the adversary.
%\jnote{Change. Compare $A$ in the 1-world to $P$ in the 0-world. Security means
%there exists a $P$
%such that for all~$A$ the following holds: $A$ in the 1-world does not do much better than
%$P$ in the 0-world. Note that there is a canonical $P$ that does best, information-theoretically.}
%\jnote{I now wonder whether we should allow test queries in both worlds.}
%
%\begin{figure}[hbtp]
%\centering
%\fpage{.6}{
%\hpagess{.5}{.45}
%{
%\experimentv{$\ExpPriv{\setprim,\distr{}{}}{A}$}\\
%$S \getsr \distr{}{}$\\
%$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
%$z \getsr A^{\TestOracle}(M,\pubaux)$\\
%if $z \in S$ then return 1\\
%return 0\\
%
%\medskip
%\experimentv{$\ExpPrivNull{\Pi, \distr{}{}}{P}$} \\
%$S \getsr \distr{}{}$\\
%%$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
%$z \getsr P()$\\
%if $z \in S$ then return 1\\
%return 0\\
%%\medskip
%%\experimentv{$\ExpWPriv{\setprim,\distr{}{}}{A}$}\\
%%$S \getsr \distr{}{}$\\
%%$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
%%$z \getsr A^{\TestOracle}(\pubaux)$\\
%%if $z \in S$ then Return 1\\
%%Return 0\\
%}
%%
%{
%\oracle{$\TestOracle(q)$}\\
%$a \gets \Qry(M,\pubaux,\privaux,q)$\\
%return~$a$
%}
%}
%\caption{Experiments for a ``one-wayness'' style privacy
%  definition.}
%\label{fig:privacy-ow}
%\end{figure}

%\jnote{Actually, I am unsure whether a one-wayness definition is independently
%interesting if we have (and can achieve) a semantic-security definition.}
%\tsnote{Depends on what information we decide is okay to leak.  A
%  classical BF gives an~$M$ that leaks a good approximation of $|S|$.
%Do we want that to be considered semantically secure?  On the other
%hand, a classical BF is one-way in the ROM, up to the min-entropy of
%the distribution over~$S$.  So that may already give an interesting
%example that fits in the gap.} \jnote{Yes. Semantic security
%is parameterized by the leakage function, and so might or might not imply
%one-wayness.}
Our semantic-secure definition is simulation-based (see
Figure~\ref{fig:privacy-ss}).
In this definition we compare the probabilities with which an attacker $A$ outputs~1
in two experiments: one ``real'' and one ``simulated.''
In the real experiment (ss-rep-1), $A$ outputs a set $S$ and is given the
public information $\pubaux$ that results from running $\Rep(S)$.\footnote{We do not give $A$ an oracle for
issuing queries; since $A$ knows the set~$S$, such an oracle would be superfluous.
\jnote{Interestingly,
if we provided a query oracle in the real world and an oracle that returns the correct
answer in the simulated world, then the definition seems to imply
correctness also.}}
In the simulated experiment (ss-rep-0), $A$ outputs $S$ and is given $\pubaux$ output
by a simulator $\Sim$ that is only given ${\sf leak}(S)$ for some (possibly randomized)
function~${\sf leak}$.
\tsnote{Need to pass state between stages if you want~$A$ to know the
  set it previously sampled, right?} \jnote{Yes, unless by convention all algorithms
  are stateful. (smiley face). Whatever you prefer is fine with me.}
For a given function ${\sf leak}$, adversary $A$, and simulator~$\Sim$,
we define the advantage measure
$\AdvPrivSS{\Pi,{\sf leak}}{A,\Sim} \bydef
\left| \Prob{\ExpPrivSSreal{\Pi}{A}=1} -
\Prob{\ExpPrivSSsim{\Pi,{\sf leak}}{A, \Sim}=1} \right|$.
Informally, a scheme $\Pi$ is secure with respect to some leakage function ${\sf leak}$
if there is a simulator $\Sim$ such that $\AdvPrivSS{\Pi,{\sf leak}}{A,\Sim}$ is small
for all~$A$. Intuitively, this means that the public portion
$\pubaux$ of the representation of a set~$S$ leaks~${\sf leak}(S)$ and nothing more.

\begin{figure}[hbtp]
\centering
\fpage{.6}{
\hpagess{.5}{.45}
{
\experimentv{$\ExpPrivSSreal{\Pi}{A}$}\\
$S \getsr A()$ \\
$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
return $A(M,\pubaux)$\\
}
%
{
\experimentv{$\ExpPrivSSsim{\Pi, {\sf leak}}{A, \Sim}$}\\
$S \getsr A()$\\
$(M,\pubaux) \getsr \Sim({\sf leak}(S))$ \\
return $A(M,\pubaux)$\\

%\medskip
%\experimentv{$\ExpPrivSSsim{\Pi}{A, \calS}$}\\
%$S \gets A()$\\
%$(M,\pubaux,\privaux) \getsr \Rep(S)$\\
%$(M',\pubaux) \gets \calS(\pubaux)$ \\
%return $A(M',\pubaux)$\\
}
}
\caption{Experiments for a ``semantic-security'' style privacy
  definition.}
\label{fig:privacy-ss}
\end{figure}

### 2016-08-29 syntax.tex
privacy definitions

>>>>

\jnote{I think we agreed to just have the syntax output $(\pub, \privaux)$, with
the semantics being that in our definitions $\pub$ is always
  public and $\privaux$ is always private.
  Different schemes can then choose whether $M$ is in $\pub$
  or $\privaux$. If we do this, then need to be careful how to define size, though
  I don't think we ever explicitly care about size for any of our results.}
